\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

This work would not have been possible without the assistance and support of a group of people who I have had the privilege to cross paths with in my journey.

Firstly I would like to express my gratitude to Professor Gérald Estadieu who, both as my supervisor and as my colleague, has had the patience to hear my ramblings about cryptocurrencies and countless project ideas for the past 15 years. For all his support, encouragement and inspiration, a huge thank you!

I would also like to thank Professor Carlos Sena Caires and Professor Álvaro Barbosa for their leadership and vision in the Faculty of Arts and Humanities, and for creating a working environment where creativity and innovation can occur, an increasingly rare commodity within our geopolitical context.

A special thanks is reserved for my great friend, and colleague, Professor José Manuel Simões, who is a truly inspiring human being and for always helping bring out the best in me. O Captain, my Captain!

I also owe a great debt of gratitude to my colleagues in the Department of Media, Art and Technology, Professors Filipa Martins, and João Brochado for all their support and above all for their patience when dealing with my absence from the office during the times I needed space to focus on this work.

I would like to thank Professor Adérito Marcos for inviting me to share my work-in-progress in the 2023 edition of the USJ Doctoral Symposium, and for his support during the restructuring of my PhD programme between faculties.

A great deal of gratitude to my colleague and friend Marco Leong, for all his valuable insights in the latest trends in technology, and for suggesting that I look into AdonisJS as a development stack.

Thank you also to Anthony Ifeanyi Okorocha, for all his support when dealing with my endless questions about the formalities of thesis submission, even during his annual leave!

I would like extend my thanks to all my other colleagues at USJ, academic, administrative and support staff, who all contribute to making this institution a workplace that I truly cherish and am proud to belong to. 

Having spent the first two decades of my life dreaming about pursuing a career in the arts, life took me in a different direction, towards the field of computing and software development. Now that I find myself veering into a new career in academia as a lecturer and researcher, I accidentally stumbled across the art world again. Within the turbulent world of crypto and web3, I found an amazing cultural oasis within the Tezos art community. It is to this community that I also dedicate this thesis. 

A special thanks to Regina Harsanyi, for pioneering the field of conservation of blockchain art, and for her invaluable insights and assistance.

A great deal of gratitude to the Teia core team, especially Zih0r for his untiring assistance whenever I had questions about IPFS and the Teia infrastructure.

Also thanks to my co-members of the hicathon's collab contract group (wg3.2), Ztepler (@ztepler9000), Steve Jones (@1x1\_music), Mical Noelson (@micalnoelson), and Shaman (@Shamanovitch) for all their amazing work.

I would like to thank all the artists who supported this work, by participating in interviews, or simply by providing encouraging feedback, namely Raphael de Courville (@sableraph), James Bloom (@crashblossom1), 
Mario Klingemann (@quasimondo), Kath O'Donnell (@kathodonnell), Thomas Lin Pedersen (@thomasp85), and Yazid (@Yazid).

Finally, I would like to thank my wife and my daughter for their patience, love and support during this journey. They gave me the strength that I needed to complete this work.


